Narrative

### 2023-05-20 manuscript/introduction.qmd
intro: improved first 3 paragraphs

Meta-analytic estimates are highly dependent on the quality of the data in the primary studies. Unfortunately, there is concern that researchers in psychology tend to neglect proper measurement (ref measurement-schmesurement) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021). Such inattention to measurement can bias heterogeneity estimates. From a statistical perspective, heterogeneity is defined as the variance between 'true' subpopulation effect sizes. However, true subpopulation effect sizes can refer to two different entities. Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as measurement error, restriction of range, and dichotomization; Hunter & Schmidt, 2015). Of these measurement artifacts, measurement error (imperfect reliability of measurement) is present in any empirical effect size estimate and systematically attenuates effect sizes. However, whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for imperfect reliability is implicit only in the models of Hunter & Schmidt (2015). This has consequences for heterogeneity estimates reported in many areas of psychology and in this article we use Monte-Carlo simulations to explore the consequences for heterogeneity estimates in meta-analysis from not correcting for unreliability.

@@

One reason that few meta-analysts correct for unreliability could be that correcting for unreliability is generally controversial and has been so since it was proposed (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017). On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., Seymour, 1988; LeBreton et al., 2017). Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1). Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas). Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated (e.g., REF1, REF2) due to factors such as publication bias (REF1) and _p_-hacking (REF2), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer. On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures. As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest. From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015). In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses should be aware of how this decision affects heterogeneity estimates.

>>>>

Meta-analytic estimates are highly dependent on the quality of the data in the primary studies that the method summarizes. However, there is concern that researchers in psychology tend to neglect proper measurement (ref measurement-schmesurement) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021). For example, Flake et al. (2017) report that although latent constructs are widely examined in personality and social psychology, researchers barely report any of the validity evidence required to ascertain the extent to which implemented scales measure the constructs of interest. Such inattention to measurement can also bias heterogeneity estimates. We will focus here on reliability of measurement (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., REF), which is the only measurement validity evidence researchers report consistently (Flake et al., 2017), and which can be corrected for in meta-analysis (Hunter & Schmidt; 2015).

Measurement error (imperfect reliability) is present in any empirical effect size estimate and systematically attenuates observed effect sizes compared to true underlying effect sizes. This affects heterogeneity estimates, which from a statistical perspective estimate the variance between true effect sizes. However, reliability is not necessarily accounted for in heterogeneity estimates, because ‘true’ underlying effect sizes can refer to two different entities. Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as imperfect reliability, restriction of range, and dichotomization; Hunter & Schmidt, 2015). Whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for measurement artifacts is implicit only in the models of Hunter & Schmidt (2015). This has consequences for heterogeneity estimates reported in many areas of psychology. Inattention to measurement reliability and how it affects heterogeneity estimates can lead researchers to misinterpret their meta-analytic average effect size estimate, ignore the presence of moderators or search for (and discover) non-existent moderators, and to implement research interventions inappropriately.

@@

One reason that few meta-analysts correct for unreliability could be that correcting for unreliability is generally controversial and has been so since it was proposed (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017). On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., Seymour, 1988; LeBreton et al., 2017). Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1). Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas). Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated (e.g., REF1, REF2) due to factors such as publication bias (REF1) and _p_-hacking (REF2), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer. On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures. As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest. From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015). In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses need to be aware of how this decision affects heterogeneity estimates.
